### 2021-07-19 chicago-spacial-covid.tex
Began report

\documentclass[10pt,letterpaper,draft]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\author{Trent Gerew}
\title{Reaction-diffusion spacial modeling of COVID-19 in Chicago}

\begin{document}
	\begin{align}
		S_t &= - \beta_{SA} SA - \beta_{SI} SI - \mu S, \\
		E_t &= \beta_{SA} SA - \beta_{SI} - (\sigma_A + \sigma_I) E, \\
		A_t &= \sigma_A E - M_{AR} A, \\
		AR_t &= M_{AR} A, \\
		I_t &= \sigma_I E - MI, \\
		H_t &= \gamma M I - (1 - \omega) \chi H - \omega \psi H, \\
		R_t &= (1 - \gamma) M I + (1 - \omega) \chi H, \\
		D_t &= \omega \psi H.
	\end{align}
\end{document}