### 2021-08-23 pres-8-27/slides/background.tex
Initial SoReMo presentation

\begin{slide}{Background}
The study of infectious diseases began with Bernoulli in 1760. \\
The study of a critical level of transmissibility for a disease is fundamental in epidemiology. \\
This is useful for
\begin{itemize}
	\item predicting counterintuitive effects of treatments, and
	\item considering containment measures
\end{itemize}
\vfill
The main flavors of models fit under deterministic/stochastic and discrete/continuous. \\

Here we consider a deterministic and continuous compartment model (Kermack-McKendrick model). \\
\begin{columns}[T]
\begin{column}{.6\textwidth}
	\begin{block}{Assumptions}
		\begin{itemize}
			\item Birth rate balances the mortality rate.
			\item No immunity in early stages of epidemic.
			\item Removal provides complete immunity against reinfection.
		\end{itemize}
	\end{block}
\end{column}
\end{columns}

\end{slide}