### 2021-08-23 pres-8-27/slides/basic_r_num.tex

\begin{slide}{Basic Reproduction Number}

Consider the disease-free equilibrium: $S=N$, $I=R=0$.\\
\vfill
Then the Jacobian matrix of the infection subsystem is
$$\mathbf{J_0}	= \begin{bmatrix} \frac{\partial \dot{I}}{\partial I} \end{bmatrix}
			= \begin{bmatrix} - \gamma + \frac{\beta_I S I}{N^2} + \frac{\beta_I S}{N} \end{bmatrix}_{S=N , I = R = 0}
			= - \gamma + \beta_I $$

$\mathbf{J_0}$ can be decomposed into transmission ($\mathbf{T} = \beta_I$) and transition ($\mathbf{\Sigma} = - \gamma$) parts such that $\mathbf{J_0} = \mathbf{T} + \mathbf{\Sigma}$. \\
\vfill
The next-generation matrix is
$$\mathbf{K_L} = - \mathbf{T} \mathbf{\Sigma}^{-1} = \frac{\beta_I}{\gamma}$$
and the basic reproduction number is
$$\mathcal{R}_0 = \rho ( \mathbf{K_L} ) = \frac{\beta_I}{\gamma}$$

\end{slide}