### 2021-08-23 pres-8-27/slides/chicago_dem.tex

\begin{slide}{Demographics and Death Rates}
The City of Chicago Data Portal provides COVID data in demographic categories for age, race, and sex. \\
We begin by fitting $\alpha$ for each category independently, where
\begin{itemize}
	\item $\delta$, $\gamma$, and $\rho$ are constants determined medical literature, and
	\item $R_0$ is determined from the Rt Live model.
\end{itemize}

\begin{columns}[T]
	\begin{column}{.4\textwidth}
		\vspace{0.5in}
		The average death rate of Chicago is then calculated by a weighted mean:
		$$\alpha = \frac{1}{\sum N_i} \sum \alpha_i N_i$$
		\vfill
	\end{column}
	\begin{column}{.5\textwidth}
		\begin{figure}[h]
			\centering
			\includegraphics[height=1.5in]{images/chicago_model}
			\caption{Final model trained on the entire data set and predicted 10 days out (from July 16).}
		\end{figure}
	\end{column}
\end{columns}

\end{slide}