### 2021-08-23 pres-8-27/slides/local_dynamics.tex

\begin{slide}{Dynamics}

\begin{columns}[T]
	\begin{column}{.7\textwidth}
		Let $k = \frac{\beta_I}{N}$, then $\dot{S} = -k S I$ and $\dot{I} = k S I - \gamma I$

		There are infintely non-isolated fixed points of the form $(a,0)$.
		Linearizing, we get
		$$\mathbf{A_{(a,0)}} = \begin{bmatrix} 0 & -k a \\ 0 & k a - \gamma \end{bmatrix}$$
		so for $a < \frac{\gamma}{k}$ the nodes are stable, and they are unstable for $a > \frac{\gamma}{k}$.
	\end{column}
	\begin{column}{.25\textwidth}
		\centering
		\includegraphics[height=1.5in]{images/local}
	\end{column}
\end{columns}
Taking $\frac{\dot{I}}{\dot{S}} = -1 + \frac{\gamma}{k S}$ and solving this differential equation gives $I = - S + \frac{\gamma}{k} \ln (S) + C$.
Thus
$$E(S,I) = S + I - \frac{\gamma}{k} \ln (S)$$
is a conserved quantity for the system.
So there cannot be any attracting fixed points (\cite{strogatz:2000}).
\end{slide}