### 2021-08-23 pres-8-27/slides/mobility.tex

\begin{slide}{Mobile Contact Probability}
If reliable mobility data exists, we can explicitly define the contact probabilities $C_{ij}^X$.
\vfill
\begin{columns}[T]
\begin{column}{.7\textwidth}
	\begin{definition}[\cite{Gatto10484}]
		$$ C_{ij}^X =
		\begin{cases}
			(1 - p_i) + (1 - r_X) p_i + r_X p_i q_{ij}	&	\text{if } i = j, \\
			r_X p_i q_{ij}					&	\text{otherwise}
		\end{cases} $$
	\end{definition}
\end{column}
\end{columns}
\vfill
\begin{itemize}
	\item $p_i$ is the fraction of mobile people in $i$
	\item $q_{ij}$ is the fraction of mobile people beween $i$ and $j=1, \dots , n$
	\item $r_X (0 \leq r_X \leq 1)$ is the fraction of contacts occuring when people of $X$ travel
\end{itemize}
\vfill
Mobility data can be often be found in the form of Origin-Destination matrices.
\begin{itemize}
	\item The \href{https://www.bts.gov/topics/airlines-and-airports/origin-and-destination-survey-data}{US DoT} and \href{https://libguides.mit.edu/gis}{MIT}
	provide such data on a national scale.
	\item Within Chicago, we can access tranportation datasets provided by the \href{https://data.cityofchicago.org/browse?category=Transportation&page=1}{Chicago Data Portal}.
\end{itemize}

\end{slide}