### 2021-08-23 pres-8-27/slides/model_contact.tex

\begin{slide}{Mobile Contact Probability}
Using one of the definitions of $\pi_{ij}$ we can define the contact probabilities.
\begin{columns}[T]
\begin{column}{.7\textwidth}
	\begin{definition}
		$$ C_{ij}^X =
		\begin{cases}
			(1 - \sum_{j=1}^n \pi_{ij}) + (1 - r_X)( \sum_{j=1}^n \pi_{ij} ) + r_X \pi_{ij} &	\text{if } i = j, \\
			r_X \pi_{ij}				&	\text{otherwise}
		\end{cases} $$
	\end{definition}
\end{column}
\end{columns}
\vfill
\begin{itemize}
	\item $r_X (0 \leq r_X \leq 1)$ is the fraction of contacts occuring when people of $X$ travel
\end{itemize}
\end{slide}