### 2021-08-23 pres-8-27/slides/spatial_model.tex

\begin{slide}{Spatial Model}

Now suppose we have a set of $n$ connected communities.
Assume $\gamma$ is community-independent, while $\lambda_i$ is the force of infection for community $i$.
\begin{columns}[T]
	\begin{column}{.5\textwidth}
		\begin{align*}
			\dot{S_i}	&=	-\lambda_i S_i \\
			\dot{I_i}	&=	\lambda_i S_i - \gamma I_i \\
			\dot{R_i}	&=	\gamma I_i
		\end{align*}
		\begin{definition}[\cite{Gatto10484}]
			$$\lambda_i = \sum_{j=1}^n C_{ij}^S \frac{ \sum_{k=1}^n ( \beta_I C_{kj}^I I_k ) }{ \sum_{k=1}^n ( C_{kj}^S S_k + C_{kj}^I I_k + C_{kj}^R R_k ) }$$
		\end{definition}
	\end{column}
	\begin{column}{.4\textwidth}
		\centering
		\includegraphics[height=1.75in]{images/map}	
	\end{column}	
\end{columns}

\vfill

$C_{ij}^X$ (with $X \in \{S,I,R\}$) is the probability that individuals who belong to compartment $X$ and are from community $i$ contact individuals in community $j$.

\end{slide}