### 2021-08-24 pres-8-27/slides/covid-01.tex
SoReMo updates

\begin{slide}{The COVID Problem}
	Identified in Wuhan, China in December 2019. \\

	Caused by the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2).

	\vspace{1cm}

	In the United States alone, we currently have
	\begin{itemize}
		\item \textbf{37,768,911} total cases
		\item \textbf{626,833} total deaths
	\end{itemize}
\end{slide}