### 2021-08-24 pres-8-27/slides/covid-03.tex

\begin{slide}{The COVID Problem}
	\centering
	\begin{columns}[T]
		\begin{column}{.5\textwidth}
			\centering
			COVID-19 Weekly Cases per 100,000 Population by Race/Ethnicity, United States
			\includegraphics[height=4.5cm]{images/ethnicity-trends}
		\end{column}
		\begin{column}{.5\textwidth}
			\centering
			COVID-19 Weekly Case Rate per 100,000 Population by Percentage of County Population in Poverty, United States
			\includegraphics[height=3.7cm]{images/poverty-trends}
		\end{column}
	\end{columns}
\end{slide}