### 2021-08-24 pres-8-27/slides/covid-01.tex
Added Covid image

\begin{slide}{The COVID Problem}
	\centering
	\begin{columns}[T]
		\begin{column}{.6\textwidth}
			Identified in Wuhan, China in December 2019. \\

			Caused by the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2).

			\vspace{1cm}

			In the United States alone, we currently have
			\begin{itemize}
				\item \textbf{37,768,911} total cases
				\item \textbf{626,833} total deaths
			\end{itemize}
		\end{column}
		\begin{column}{.2\textwidth}
			%\centering
			\includegraphics[height=4cm]{images/covid}
		\end{column}
	\end{columns}
\end{slide}