### 2021-08-26 pres-8-27/slides/work-01.tex
SoReMo updates

\begin{slide}{Current Work}
	Working over the summer with Prof. Chun Liu and Prof. Yiwei Wang. \\

	\begin{enumerate}
		\item Start with a population of \emph{susceptibles} ($S$)
		\item Some of the population may become \emph{infected} ($I$) upon emergence of the virus
		\item Infected individuals interact with susceptibles at rate $\beta$ to draw new members into $I$
	\end{enumerate}
\end{slide}