### 2021-08-26 pres-8-27/slides/work-03.tex
SoReMo updates

\begin{slide}{Current Work}
	{\large Next we identify the parameters $\beta$ and $\gamma$ at the ODE level:} \\

	\vspace{.5cm}

	\textbf{Idea:} Minimize the distance between the model time series and the observed time series.
	\begin{equation*}
	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| C_\num (t_i) - C_\obs (t_i) \right|^2 + \left| D_\num (t_i) - D_\obs (t_i) \right|^2 \right)
	\end{equation*}
	where the index $i$ identifies a point in the time series.\\

	\vspace{.5cm}

	Optimize the parameters to reproduce:
	\begin{itemize}
		\item the reported \emph{total} number of infected cases $C(t) = I(t) + R(t)$
		\item the \emph{total} number of deceased $D(t) = R(t)$.
	\end{itemize}

\end{slide}