### 2021-08-26 pres-8-27/slides/social-01.tex
SoReMo updates

\begin{slide}{Social and Community Connections}
	{\large Original context:} Working with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\

	\vspace{.2cm}

	What are the effects of \textbf{age} and \textbf{race/ethnicity} in the Chicago area?
\end{slide}