### 2021-08-27 pres-8-27/slides/work-00.tex
SoReMo updates

\begin{slide}{Current Work}
	{\large \textbf{Idea:}}
	\begin{enumerate}
		\item Define a \textit{local} process that occurs locally at every point in space.
		\item Determine suitable \textit{local} parameters (e.g., transition rates).
		\item Define the characteristics of the \textit{physical space} (e.g., population density).
		\item Integrate the local process over the physical space.
		\item Determine suitable \textit{physical} parameters (e.g., transmission rates).
	\end{enumerate}

	\vspace{.5cm}

	{\large \textbf{Progress:}}
	\begin{itemize}
		\item Start by defining a toy model: \textbf{S}usceptible, \textbf{I}nfected, \textbf{R}emoved.
		\item Local parameters fitted using data from the \alert{Chicago Data Portal}.
		\item Spatial characteristics defined using data from the \alert{U.S. Census Bureau}.
		\item Numerical spatial integration is a work in progress.
	\end{itemize}
\end{slide}