### 2021-10-26 pres-8-27/Plaincolortheme.tex
Lots of updates.

% This effectively sets an IIT color theme with no background bars and is
% derived from the beaver theme that is built into beamer
% 
\definecolor{iitred}{rgb}{0.7423,0.0020,0.0192}

\setbeamercolor{section in toc}{fg=black,bg=white}
\setbeamercolor{alerted text}{fg=iitred}
\setbeamercolor*{palette primary}{fg=iitred,bg=white}
\setbeamercolor*{palette secondary}{fg=iitred,bg=white}
\setbeamercolor*{palette tertiary}{bg=white,fg=iitred}
\setbeamercolor*{palette quaternary}{fg=iitred,bg=white}

\setbeamercolor*{sidebar}{fg=iitred,bg=white}

\setbeamercolor*{palette sidebar primary}{fg=iitred}
\setbeamercolor*{palette sidebar secondary}{fg=iitred}
\setbeamercolor*{palette sidebar tertiary}{fg=iitred}
\setbeamercolor*{palette sidebar quaternary}{fg=iitred}

%\setbeamercolor*{titlelike}{parent=palette primary}
\setbeamercolor{titlelike}{parent=palette primary,fg=iitred}
\setbeamercolor{frametitle}{bg=white}
\setbeamercolor{frametitle right}{bg=white}

\setbeamercolor*{separation line}{}
\setbeamercolor*{fine separation line}{}
\setbeamercolor{block body}{parent=normal text,use=block title,bg=red!0,fg=}
\setbeamercolor{local structure}{fg=iitred}

% This sets all text color
%\setbeamercolor{normal text}{fg=iitred}