### 2021-10-26 pres-8-27/slides/future-02.tex

\begin{slide}{Future Goals}
	{\large Transition to PDE model (proof of concept)}

	\vspace{.5cm}

	\begin{itemize}
		\item The transition rate $\gamma$ is an ``individual'' parameter, so the value is preserved at the PDE level.
		\item The transmission rate $\beta$ depends on the interaction between individuals, so this must be modified.

		\vspace{.3cm}

		\item The geometry (shape and density distribution) of Chicago must be defined (Census data).
		\item Suitable initial infected distributions must be seeded.
	\end{itemize}
\end{slide}