### 2021-10-26 pres-8-27/slides/social-01.tex

\begin{slide}{Social and Community Connections}
	{\large Summer 2020 context:} IPRO with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\

	\vspace{.2cm}

	\begin{itemize}
		\item How does the impact of COVID-19 (in Chicago) depend on \textbf{age} and \textbf{race/ethnicity}?
	\end{itemize}

	\vspace{1cm}

	{\large Present context:} COVID-19 is still here. Where is it impacting and why?

	\vspace{.2cm}

	\begin{itemize}
		\item Does \textbf{community structure} play a role?
		\item How effective are \textbf{vaccines}? (Or what is thwarting them?)
		\item Are \textbf{care facilities} properly allocated?
		\item \emph{Other ideas?}
	\end{itemize}

	\vspace{.7cm}

	{\large \textbf{Goal:} Can we design a model to answer these questions?}
\end{slide}