### 2021-10-26 pres-8-27/slides/work-02.tex

\begin{slide}{Current Work}
	The population model at the PDE level is an autonomous diffusion with a source:
	\begin{align*}
	S_t	&=	\nabla (\D_s \nabla S) - \beta SI,	\\
	I_t	&=	\nabla (\D_I \nabla I) + \beta SI - \gamma I,	\\
	R_t	&=	\gamma I.
	\end{align*}

	\vspace{.5cm}
	Assumptions:
	\begin{itemize}
		\item The $R$ population has immunity.
		\item Diffusion performs all relevant spreading.
		\item Arriving infected individuals form local hotspots.
		\item Disease progression is the only transition mechanism.
	\end{itemize}

\end{slide}