### 2021-10-26 pres-8-27/slides/work-05.tex

\begin{slide}{Current Work}
	{\large ODE model for Chicago with fitting to official data from March 17, 2020 ($t = t_\text{init} = 1$) to September 30, 2020 ($t = t_\text{fit}^\text{end} = 198$)}
	\begin{figure}
		\centering
		\subfloat[Confirmed cases $C(t) = I(t) + R(t)$]{\includegraphics[width=.333\textwidth]{images/ode_cases}}
		\subfloat[Number of deaths $D(t)$]{\includegraphics[width=.333\textwidth]{images/ode_deaths}}
		\subfloat[Infected $I(t)$ and Recovered $R(t)$]{\includegraphics[width=.333\textwidth]{images/ode_IR}}
	\end{figure}
\end{slide}